Abstract Binding Trees

### 2015-11-01 paper.tex
Add ends and Kan extensions to agda code
Define substitution tensor in Agda
Fix part about replacing multiplication
Fix up rendering
notation changes; add symbols to interp; tentative interp of metavars

% fonts
\usepackage[
  activate={true,nocompatibility},
  final,
  tracking=true
]{microtype}
\setmainfont{Linux Libertine O}
\setmathfont{Asana Math}

@@

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\langle #1, #2 \rangle}
\newcommand\Tuple[1]{\langle #1 \rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle
}
\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

@@

Then, variables object
$\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid \Gamma(x)=s}}$
is the unit to this tensor. We will say that an object
$\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
with the following natural transformations where $\nu$ embeds variables into $P$
and $\varsigma$ equips $P$ with an operation for simultaneous substitutions of
variables.

@@

\begin{remark}
  A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
  also forms a \emph{relative monad}\footnote{Relative monads are skew-monoids
    in the skew-monoidal structure induced by a functor category
    $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. However,
    relative monads can also be understood without reference to monoids by
    viewing them as a generalization of \emph{extension systems} or
    \emph{Kleisli structures}. This latter presentation of monads, popular in
    computer science and logic, does not involve iteration of the monad
    functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
  Consequently, the monoid multiplication $\odot$ can be replaced with an
  extension operation:
  \[
    \begin{tikzcd}[ampersand replacement = \&, sep = huge]
      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{\circledast}"]
      \& {\IMode{P(\,=\,)^{P(\,-\,)}}}
    \end{tikzcd}
  \]
\end{remark}

For any context $\Gamma$, we have an induced map
$\Of{\nu_\Gamma}{\left(\prod_{x\in\Dom\Gamma}V_{\Gamma(x)}\right)\to
\left(\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}\right)}$.

@@

\subsection{Interpretations of terms}

\[
  \Define{
    \SemBrackets{
      \Omega\triangleright\Upsilon\parallel\Gamma
    }_P
  }{
    \left(\prod_{(\MV{m}:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}
    P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}})}\right)
    \times
    \prod_{x\in\Dom\Gamma}V_{\Gamma(x)}
  } %% ???
\]

Let $\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$ be
defined as follows:
\begin{align*}
  \ADefine{
    \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j : s}_P
  }{
    \begin{tikzcd}[ampersand replacement = \&]
      {\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P} \arrow[r, "\pi_2"]
      \& { \prod_{x\in\Dom\Gamma}V_{\Gamma(x)} } \arrow[r, "\nu_{\Gamma}"]
      \& {\prod_{x\in\Dom\Gamma} P_{\Gamma(x)}} \arrow[r, "\pi_j"]
      \& P_s
    \end{tikzcd}
  }
\end{align*}

@@

\subsection{Nat}

\ExecuteMetaData[latex/Code.tex]{nat}

\subsection{Fin}

\ExecuteMetaData[latex/Code.tex]{fin}

\subsection{Name}

\ExecuteMetaData[latex/Code.tex]{name}

\subsection{Ctx}

\ExecuteMetaData[latex/Code.tex]{ctx}

@@

>>>>

% fonts
\usepackage[
  activate={true,nocompatibility},
  draft,
  tracking=true
]{microtype}
\setmainfont{Linux Libertine O}
\setmathfont{Asana Math}

@@

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\left\langle #1, #2 \right\rangle}
\newcommand\Tuple[1]{\left\langle #1 \right\rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle
}
\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

@@

Then, the variables object
$\Define{V_s(\Upsilon\parallel\Gamma)}{\MkSet{x\in\Dom\Gamma\mid \Gamma(x)=s}}$
is the unit to this tensor; additionally, we have a presheaf of symbols
$\Define{Y_s(\Upsilon\parallel\Gamma)}{\MkSet{u\in\Dom\Upsilon\mid
\Upsilon(x)=s}}$. We will say that an object
$\Of{P}{\widehat{\HypCat}^\Sorts}$ is a $\Sigma$-monoid in case it is equipped
with the following natural transformations where $\nu$ embeds variables into
$P$ and $\varsigma$ equips $P$ with an operation for simultaneous substitutions
of variables.

@@

For an object $\Of{P}{\widehat{\HypCat}^\Sorts}$, we will use the notation
$P^{[\Gamma]}$ to mean $\prod_{x\in\Dom\Gamma}P_{\Gamma(x)}$; likewise,
$Y^{\{\Upsilon\}}$ shall mean $\prod_{u\in\Dom\Upsilon}Y_{\Upsilon(u)}$.
Additionally, we have an induced map $\Of{\nu_\Gamma}{V^{[\Gamma]}\to
P^{[\Gamma]}}$.

\begin{remark}
  A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
  also forms a \emph{relative monad}\footnote{Relative monads are skew-monoids
    in the skew-monoidal structure induced by a functor category
    $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. However,
    relative monads can also be understood without reference to monoids by
    viewing them as a generalization of \emph{extension systems} or
    \emph{Kleisli structures}. This latter presentation of monads, popular in
    computer science and logic, does not involve iteration of the monad
    functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
  Consequently, the monoid multiplication $\varsigma$ can be replaced with an
  extension operation:
  \[
    \begin{tikzcd}[ampersand replacement = \&, sep = huge]
      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{\circledast}"]
      \& {\IMode{P(\,=\,)^{P(\,-\,)}}}
    \end{tikzcd}
  \]
\end{remark}

@@

\subsection{Interpretation of terms}

\[
  \Define{
    \SemBrackets{
      \Omega\triangleright\Upsilon\parallel\Gamma
    }_P
  }{
    \left(\prod_{(\MV{m}:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}
    P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}})}\right)
    \times
    Y^{\{\Upsilon\}}
    \times
    V^{[\Gamma]}
  }
\]

We define the interpretation of a term $M$ in the model $P$ as a morphism
$\Of{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
M:s}_P}{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P\to P_s}$.
The interpretation of a variable
$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash x_j:s}_P$ is
the following composite:

\[
  \begin{tikzcd}[ampersand replacement = \&]
    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\pi_3}"]
    \& {\IMode{V^{[\Gamma]}}} \arrow[r, "\IMode{\nu_{\Gamma}}"]
    \& {\IMode{P^{[\Gamma]}}} \arrow[r, "\IMode{\pi_j}"]
    \& \IMode{P_s}
  \end{tikzcd}
  \tag{var}
\]

The interpretation of metavariable applications
$\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash
\MApp{\MV{m}_i}{u_0,\dots,u_m}{M_0,\dots,M_n} : s}_P$, where
$\Lookup{\Omega}{\MV{m}}{\MkValence{\vec{p}}{\vec{q}}{s}}$, is the following composite
\[
  \begin{tikzcd}[ampersand replacement = \&, sep=large]
    {\IMode{\SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma}_P}} \arrow[r, "\IMode{\Tuple{\pi_i\,\pi_1, \phi}}"]
    \& {\IMode{P_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}} \times P^{[\vec{q}]}}} \arrow[r, "\IMode{\varsigma_{\vec{q}}}"]
    \& \IMode{P_s}
  \end{tikzcd}
  \tag{metavar}
\]
where
\[
  \Define{\phi}{
    \Tuple{
      \SemBrackets{\Omega\triangleright\Upsilon\parallel\Gamma\vdash M_j:q_j}_P
    }_{j\leq n}
  }
\]

@@

\subsection{Kan Extensions}

\subsubsection{Lan}

\ExecuteMetaData[latex/Code.tex]{lang}\\

\ExecuteMetaData[latex/Code.tex]{lan}

\subsubsection{Ran}

\ExecuteMetaData[latex/Code.tex]{rang}\\

\ExecuteMetaData[latex/Code.tex]{ran}

\subsection{Variables and Symbols}

\ExecuteMetaData[latex/Code.tex]{var}\\

\ExecuteMetaData[latex/Code.tex]{sym}

\subsection{Contexts}

\ExecuteMetaData[latex/Code.tex]{ctx}\\

\ExecuteMetaData[latex/Code.tex]{sctx}\\

\ExecuteMetaData[latex/Code.tex]{ctxdom}\\

\ExecuteMetaData[latex/Code.tex]{ctxidx}\\

\ExecuteMetaData[latex/Code.tex]{elem}

@@

\subsection{Substitution Monoidal Structure}

\ExecuteMetaData[latex/Code.tex]{H}\\

\ExecuteMetaData[latex/Code.tex]{HHat}\\

\ExecuteMetaData[latex/Code.tex]{V}\\

\ExecuteMetaData[latex/Code.tex]{tensor0}\\

\ExecuteMetaData[latex/Code.tex]{tensor1}\\

\ExecuteMetaData[latex/Code.tex]{extension}
